# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r }
#read table, process date column, sum total steps per day
data<-read.csv("activity.csv",header=T,colClasses=c("numeric","character","numeric"))
data$date<-as.Date(data$date,format="%Y-%m-%d",tz="")
data<-data[order(data$date, data$interval),]
str(data)
totalStepsPerDay<-aggregate(steps ~ date, data, sum)
```

## What is mean total number of steps taken per day?

```{r }
#create histogram of total number of steps per day
hist(totalStepsPerDay$steps,breaks=10,
     main="Total number of steps per day",
     xlab="Steps per day")
mean.steps<-round(mean(totalStepsPerDay$steps),2)
median.steps<-round(median(totalStepsPerDay$steps),2)
```
**The mean and median for the total number of steps per day:**  
- **`r paste("Mean:", mean.steps)`**
- **`r paste("Median:", median.steps)`**

## What is the average daily activity pattern?
```{r }
averageStepsPerInterval<-aggregate(steps ~ interval, data, mean)
plot(averageStepsPerInterval$interval,averageStepsPerInterval$steps,type="l",
     xlab="Interval", ylab="Mean of steps per interval",
     main="Average number of steps per interval")
i<-which(averageStepsPerInterval$steps==max(averageStepsPerInterval$steps))
max.steps<-round(max(averageStepsPerInterval$steps),0)
av.spi<-averageStepsPerInterval$interval[i]
```
The **`r i`<sup>th</sup> interval (`r av.spi`)** has the maximum number of steps (**`r max.steps`**) on the average.

## Imputing missing values
```{r }
#inputting missing values in a new dataset(data.replacena) with means for interval
data.replacena<-data
l<-nrow(data.replacena)
for(j in 1:l){
  if(is.na(data.replacena$steps[j])){
    interval<-data.replacena$interval[j]
    nrSteps<-averageStepsPerInterval$steps[averageStepsPerInterval$interval==interval]
    data.replacena$steps[j]<-nrSteps
  }
}
str(data.replacena)
totalStepsPerDayCor<-aggregate(steps ~ date, data=data.replacena, sum)
hist(totalStepsPerDayCor$steps,breaks=10,
     main="Total number of steps per day (NAs replaced with mean for interval)",
     xlab="Steps per day")
mean2.steps<-round(mean(totalStepsPerDayCor$steps),2)
median2.steps<-round(median(totalStepsPerDayCor$steps),2)
```
**The mean and median for the total number of steps per day (when NAs replaced):**  
- **`r paste("Mean:", mean2.steps)`**
- **`r paste("Median:", median2.steps)`**  

**The mean has not changed with NAs replacements, but the median has changed** 

## Are there differences in activity patterns between weekdays and weekends?
```{r }
data.replacena$weekday<-as.factor(weekdays(data.replacena$date))
weekend.subset <- subset(data.replacena, weekday %in% c("Saturday","Sunday"))
weekday.subset <- subset(data.replacena, !weekday %in% c("Saturday","Sunday"))

averageStepsPerIntervalWE<-aggregate(steps ~ interval, weekend.subset, mean)
averageStepsPerIntervalWD<-aggregate(steps ~ interval, weekday.subset, mean)

par(mfrow=c(2,1))
plot(averageStepsPerIntervalWD$interval,averageStepsPerIntervalWD$steps,type="l",
     main="The average number of steps per interval in weekdays",
     xlab="Interval",ylab="Average number of steps")
plot(averageStepsPerIntervalWE$interval,averageStepsPerIntervalWE$steps,type="l",
     main="Average number of steps per interval in weekends",
     xlab="Interval",ylab="Average number of steps")

averageStepsPerDayWE<-aggregate(steps ~ date, weekend.subset, sum)
mean.we<-round(mean(averageStepsPerDayWE$steps),0)
averageStepsPerDayWD<-aggregate(steps ~ date, weekday.subset, sum)
mean.wd<-round(mean(averageStepsPerDayWD$steps),0)
```
**There is a greater number of steps on average in weekends days than in weekdays days**  
- **`r paste("Weekend day:", mean.we)` steps**
- **`r paste("Weekday day:", mean.wd)` steps**